\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-05 Limber_Comment.tex
photoz.tex
Accepted HH's comment (but as a footnote)
ready to send to the wider group...

\begin{abstract}
Kitching et al. (2016) examine the impact of the flat-sky and Limber approximations that can be used when constraining cosmological parameters from the measurement of weak gravitational lensing by large-scale structures, also known as cosmic shear.  We support the core message of this paper that the use of such approximations should be considered in the future analyses of high precision surveys such as Euclid, LSST and WFIRST.  In this comment, we argue that these approximations have negligible impact for current surveys and are not able to explain away the tension between the constraints from some current weak lensing surveys and Planck.  Reducing uncertainty and characterising bias in photometric redshift estimation is a critical task for the success of current and future surveys.  We highlight that sophisticated treatments of these uncertainties have already been explored in the literature and have been shown not to resolve the tension with Planck, in contrast to the message that Kitching et al. (2016) convey}.

@@

This comment is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We show that we are unable to replicate key results from \citet{kitching/etal:2016} and therefore do not expect these approximations to impact on the cosmological analysis of current surveys. In section~\ref{sec:cfhtlens} we compare with \citet{kitching/etal:2016} our conclusions of a `one-parameter' cosmological re-analysis of 2D cosmic shear data from \citet{kilbinger/etal:2013}.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published and discuss the impact in our analysis of numerical simulations and CMB lensing.

>>>>

\begin{abstract}
Kitching et al. (2016) examine the impact of the flat-sky and Limber approximations that can be used when constraining cosmological parameters from the measurement of weak gravitational lensing by large-scale structures, also known as cosmic shear.  We support the core message of this paper that the use of such approximations should be considered in the future analyses of high precision surveys such as Euclid, LSST and WFIRST.  In this comment, we argue that these approximations have negligible impact for current surveys and are not able to explain away the tension between the constraints from some current weak lensing surveys and Planck.  Reducing uncertainty and characterising bias in photometric redshift estimation is a critical task for the success of current and future surveys.  We highlight that sophisticated treatments of these uncertainties have already been explored in the literature and have been shown not to resolve the tension with Planck, in contrast to the message that Kitching et al. (2016) convey.

@@

This comment is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We find that we are unable to replicate key results from \citet{kitching/etal:2016} and show that do not expect these approximations to impact on the cosmological analysis of current surveys. In section~\ref{sec:cfhtlens} we compare with \citet{kitching/etal:2016} our conclusions of a `one-parameter' cosmological re-analysis of 2D cosmic shear data from \citet{kilbinger/etal:2013}.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published and discuss the impact of flat-sky and Limber approximations in our analysis of numerical simulations and CMB lensing.

### 2016-12-05 cfhtlens.tex
ready to send to the wider group...

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this artificial reduction in the error by a factor of $\sim 2$, in comparison to a full six-parameter fit, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} finding the lowest conditional value of $\sigma_8 = 0.743 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.744 \pm 0.023$ for the `Baseline Limber Flat` (LF) case.  As expected from the comparison in Figure~\ref{fig:Cl_xi} the different combinations of assuming flat or spherical sky, or baseline of extended Limber make little difference to the cosmological parameter constraints, even in this `one-parameter' analysis.

@@

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range different two-point shear statistics including the shear correlation function, $\xi_\pm$, the mass aperture statistic, $M_ {\rm ap}^2$ \citep{schneider/etal:1998}, and the COSEBIs statistics \citep{schneider/etal:2010}.  Figure~\ref{fig:filters} shows the integrands of these statistics for two cases,
normalised to their maximum value.
The top plot shows the integrand of $\xi_+$ for $\theta=100'$ and $\theta=350'$. 
Note that the bottom left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum
angular separation of $100'$. However, in \citet{kilbinger/etal:2013} which is the same data used in \citet{kitching/etal:2016} 
the maximum $\theta$ is $350'$. The middle panel in Figure~\ref{fig:filters} shows COSEBIs integrands for two angular ranges,
$[1',100']$ and $[0.8',350']$ which corresponds to the largest angular range used in \citet{kilbinger/etal:2013}. Here we only show the integrands for the lowest COSEBIs mode, as the higher modes generally probe larger $\ell$-modes.
Finally, the bottom panel shows the integrands of aperture mass dispersion statistics, for two maximum angular ranges, discussed above. 
As shown by each statistic's weight function in figure~\ref{fig:filters}, these two-point cosmic shear statistics exhibit different dependences between the angular scales sampled and the $\ell$-range probed.   
If a significant bias had been introduced at low-$\ell$ by using flat-sky and Limber approximations, we would then expect to see a systematic shift between the different two-point statistics with the COSEBIs statistic being essentially unaffected as it only includes modes with $\ell \gtrsim 80$.  This is found not to be the case with all three statistics finding $\sigma_8 (\Omega_m/0.27)^\alpha = 0.79$ with errors that range from $0.03$ to $0.06$ \citep[see Table 5 of][]{kilbinger/etal:2013}.  This comparison further supports our argument that the approximations highlighted by \citet{kitching/etal:2016} have negligible impact for current surveys.

@@

\subsection{On the application to CMB lensing}
\citet{kitching/etal:2016} argue that their correction to the Limber and flat sky approximations would also affect predictions involving the shear and convergence cross-spectrum, such that CMB lensing - galaxy lensing cross-correlation should include a pre-term $[\ell/(\ell+0.5)]^4$.   We have verified that this term would have negligible impact on the current measurements, given the relatively low signal to noise ratio. For example, most measurements \citep{hand/etal:2015, liu/hill:2015, kirk/etal:2016,harnois-deraps/etal:2016} are carried over angular multipoles in the range $\ell \in [20-2000]$, and are organized in $\sim$ five bins. The size of the pre-term $[\ell/(\ell+0.5)]^4$, averaged over the first bin, is of the order of 2\%, and then becomes sub-percent for all other bins. In comparison, the size of the statistical error bars are of the order 25-35\% with additional uncertainty due to photometric redshifts (10-15\%) and intrinsic galaxy alignments (10-15\%).

>>>>

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this artificial reduction in the error by a factor of $\sim 2$, in comparison to a full six-parameter fit, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} finding the lowest conditional value of $\sigma_8 = 0.743 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.744 \pm 0.023$ for the `Baseline Limber Flat` (LF) case.  As expected from the comparison in Figure~\ref{fig:Cl_xi} the different combinations of assuming flat or spherical sky, baseline or extended Limber, make little difference to the cosmological parameter constraints, even in this `one-parameter' analysis.

@@

We note that it appears \citet{kitching/etal:2016} have chosen to use their `Flat' approximation in their `Full' non-Limber analysis rather than a full spherical-sky solution.  The rationale is based on the Hankel transform in equation~\ref{eqn:xiGG} which is derived assuming a flat-sky approximation.  As shown in Figure~\ref{fig:Cl_xi}, however their chosen version of a flat-sky approximation serves to exacerbates the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution. We do not find this to be the case when considering the Standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range different two-point shear statistics including the shear correlation function, $\xi_\pm$, the mass aperture statistic, $M_ {\rm ap}^2$ \citep{schneider/etal:1998}, and the COSEBIs statistics \citep{schneider/etal:2010}.  Figure~\ref{fig:filters} shows the integrands of these statistics (i.e. the weights placed on each $\ell$ scale),  for two cases, normalised to their maximum value.  The upper panel shows the integrand\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum
angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, 
the maximum $\theta$ is $350'$.} of $\xi_+$ for $\theta=100'$ and $\theta=350'$. The middle panel in Figure~\ref{fig:filters} shows COSEBIs integrands for two angular ranges,$[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, as the higher modes generally probe larger $\ell$-modes.  Finally, the lower panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.
From Figure~\ref{fig:filters} we can see that the two-point cosmic shear statistics tested by \citet{kilbinger/etal:2013} exhibit different dependences between the angular scales sampled and the $\ell$-range probed.   
If a significant bias had been introduced at low-$\ell$ by using flat-sky and Limber approximations, we would then expect to see a systematic shift between the different two-point statistics with the COSEBIs statistic being essentially unaffected as it only includes modes with $\ell \gtrsim 80$.  This is found not to be the case with all three statistics finding $\sigma_8 (\Omega_m/0.27)^\alpha = 0.79$ with errors that range from $0.03$ to $0.06$ \citep[see Table 5 of][]{kilbinger/etal:2013}.  This comparison further supports our argument that the approximations highlighted by \citet{kitching/etal:2016} have negligible impact for current surveys.

@@

\subsection{On the application to CMB lensing}
\citet{kitching/etal:2016} argue that their correction to the Limber and flat sky approximations would also affect predictions involving the shear and convergence cross-spectrum, such that CMB lensing - galaxy lensing cross-correlation should include a pre-term $ [(\ell*(\ell+1))^2]/(\ell+1/2)^4.$.   We have verified that this term would have negligible impact on the current measurements, given the relatively low signal to noise ratio. For example, most measurements \citep{hand/etal:2015, liu/hill:2015, kirk/etal:2016,harnois-deraps/etal:2016} are carried over angular multipoles in the range $\ell \in [20-2000]$, and are organized in $\sim$ five bins. The size of the pre-term $[\ell/(\ell+0.5)]^4$, averaged over the first bin, is of the order of 2\%, and then becomes sub-percent for all other bins. In comparison, the size of the statistical error bars are of the order 25-35\% with additional uncertainty due to photometric redshifts (10-15\%) and intrinsic galaxy alignments (10-15\%).

### 2016-12-05 theory.tex
ready to send to the wider group...

As shown in the lower panels of Figure~\ref{fig:Cl_xi}, the `standard flat' sky approximation used by recent cosmic shear surveys, where $T_\ell = 1$, recovers a very similar result to the spherical sky case.  Over angular scales used in the recent KiDS analysis (shown dashed), these curves differ by less than 1\% and such the standard flat-sky approximations that have been used will not impact the cosmological analyses of current surveys.

>>>>

As shown in the lower panels of Figure~\ref{fig:Cl_xi}, the `standard flat' sky approximation used by recent cosmic shear surveys, where $T_\ell = 1$, recovers a very similar result to the spherical sky case.  Over angular scales used in the recent KiDS analysis (shown dashed), these curves differ by less than 1\% as such the standard flat-sky approximations that have been used will not impact the cosmological analyses of current surveys.
